---
title: "Exporatory analysis grocery"
author: "Erika"
format: html
editor: visual
---

## About the data

This is a study about the data collected from receipts of daily grocery shop in for a couple in Ireland. The main purpose of this analysis is to support the construction of an app that register and analyse this kind of data.

```{r}
#| echo: false
#| include: false
library(dplyr)
library(tidyr)
library(readr)
library(echarts4r)
library(stringr)
library(lubridate)
library(purrr)

grocery_table <- read_rds("./Data/grocery_table.rds")
product_table <- read_rds("./Data/product_table.rds")

week_of_month <- function(date) {
  first_day <- floor_date(date, "month")
  (day(date) + wday(first_day) - 1) %/% 7 + 1
}

grocery <-  grocery_table |>
  left_join(product_table) |>
  mutate(Year = grocery_day |> year(),
         Month = grocery_day |> month(),
         Week_of_month = week_of_month(grocery_day),
         Weekday = grocery_day |> wday()
         )



```

### Grocery day

Grocery days for each year

```{r}
#| message: false
#| warning: false
#| comment: false
grocery_by_day <- grocery |>
  group_by(grocery_day) |>
  summarise(total_purchase = value_total |> 
              sum(na.rm = TRUE)) 
grocery_by_day |>
  e_chart(grocery_day) |>
  e_bar(total_purchase) |>  
  e_title("Total purchase by date")
  

```

Histogram purchase by day

```{r}
grocery_by_day |>
  e_charts(total_purchase) |>
  e_histogram(total_purchase, name = "Histogram") |>
  e_density(total_purchase, name = "Density", areaStyle = list(opacity = .4), smooth = TRUE, y_index = 1) |>
  e_tooltip(trigger = "axis")
```

Boxplot grocery by day

```{r}
 grocery_by_day |>
   e_charts() |>
   e_boxplot(total_purchase) |>
  e_title("Distribution by day") |> 
  e_tooltip(trigger = "axis" )
   
```

Value box grocery by month, Year

```{r}
grocery_by_day |>
  mutate(Year = grocery_day |> year(),
         Month = grocery_day |> month(),
         year_month = )
```

Average value by purchase for each year

```{r}
grocery |>
  group_by(grocery_day, Year) |>
  summarise(total_purchase = value_total |> 
              sum(na.rm = TRUE)) |>
  ungroup() |>
  group_by(Year) |>
  summarise(mean_purchase = total_purchase |> mean()) |>
  ungroup() |>
  mutate(Year = Year |> as.character()) |>
  e_chart(Year) |>
  e_bar(mean_purchase) |>
  e_title("Mean purchase by Year") |> 
  e_tooltip(trigger = "axis" )
```

Total spent by month and Year

```{r}
grocery |>
  group_by(Year, Month) |>
  summarise(total_grocery = value_total |> 
              sum(na.rm = TRUE)) |>
  group_by(Year) |>
  e_charts(Month) |>
  e_line(total_grocery) |>
  e_title("Total spent in a month") |>
  e_tooltip(trigger = "axis")
  
  
```

```{r}
grocery |>
  group_by(Year, Month) |>
  summarise(total_grocery = value_total |> 
              sum(na.rm = TRUE)) |>
  group_by(Year) |>
e_charts(Month) |>
  e_bar(total_grocery) |>
  e_title("Total spent in a month") |>
  e_tooltip(trigger = "axis")
```

How often do groceries by month Year

```{r}
grocery |>
  distinct(Year, Month, grocery_day) |>
  group_by(Year, Month) |>
  summarise(count_groceries = n()) |>
  group_by(Year) |>
  e_charts(Month) |>
  e_bar(count_groceries) |>
  e_title("Total groceries in a month") |>
  e_tooltip(trigger = "axis")
  
```

Frequency of groceries by day of the week

```{r}
grocery |>
  group_by(grocery_day, Weekday) |>
  summarise(total_value = value_total |> sum(na.rm = TRUE)) |>
  ungroup() |>
  group_by(Weekday) |>
  summarise(frequency_weekday = n(),
            avg_value = total_value |> mean()) |>
  e_charts(Weekday) |>
  e_bar(frequency_weekday, name = "Weekday frequency") |>
  e_line(avg_value,name = "Average spent", y_index = 1) |>
  e_title("Total groceries in a week") |>
  e_tooltip(trigger = "axis")


```

Frequency by week number

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n()) |>
  ungroup() |>
  group_by(Year) |>
  summarise(avg_count_groc = count_groceries |> mean())
```

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n() ) |>
            ungroup() |>
  mutate(year_month = paste(Year, Month, Week_of_month, sep = "_")) |>
    e_charts(year_month) |> 
  e_line(count_groceries) |> 
  e_title("Groceries per week")  |>
  e_tooltip(trigger = "axis")
  
```

frequency by week number

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n() ) |>
  ungroup() |>
  group_by(Week_of_month) |>
  summarise(avg_week = count_groceries |> sum())  |> 
  e_charts(Week_of_month) |> 
  e_bar(avg_week) |> 
  e_tooltip("axis")

```

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n()) |>
  ungroup() |>
  group_by(Year, Week_of_month) |>
  summarise(frequency_week_of_month = count_groceries |> sum())
  
```

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n()) |>
  ungroup() |>
  group_by(Year, Month) |>
  summarise(avg_count_groceries_by_week = count_groceries |> mean()) |>
  ungroup() |>
  mutate(year_month = paste(Year, Month, sep = "_")) |>
    e_charts(year_month) |> 
  e_line(avg_count_groceries_by_week) |> 
  e_title("Groceries per week")  |>
  e_tooltip(trigger = "axis")
```

```{r}
grocery |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n()) |>
  ungroup() |>
  group_by(Year, Month) |>
  summarise(count_groceries_by_month = count_groceries |> sum()) |>
  ungroup() |>
  mutate(year_month = paste(Year, Month, sep = "_")) |>
    e_charts(year_month) |> 
  e_line(count_groceries_by_month) |> 
  e_title("Groceries per month")  |>
  e_tooltip(trigger = "axis")
```

```{r}
supermakets_choosen <- c('Tesco', 'Lidl', 'Aldi')

grocery |>
  filter(supermarket_name %in% supermakets_choosen) |>
  distinct(Year, Month, Week_of_month, grocery_day) |>
  group_by(Year, Month, Week_of_month) |>
  summarise(count_groceries = n()) |>
  ungroup() |>
  group_by(Year, Month) |>
  summarise(count_groceries_by_month = count_groceries |> sum()) |>
  ungroup() |>
  mutate(year_month = paste(Year, Month, sep = "_")) |>
    e_charts(year_month) |> 
  e_line(count_groceries_by_month) |> 
  e_title("Groceries per month")  |>
  e_tooltip(trigger = "axis")
```

Calendar groceries

```{r}
data_calendar <- grocery |>
  group_by(grocery_day, Year) |>
  summarise(grocery_value = value_total |> 
              sum(na.rm = TRUE)) |>
  arrange(grocery_day, Year)

calendar_plot <- function(data, year_choose){
  data2 = data |> 
  filter(Year == year_choose)
  
  max_value = data2$grocery_value |> max()
  
  data2 |>
  group_by(Year) |> 
  e_charts(grocery_day) |> 
  e_calendar(range = year_choose, top="40") |> 
  e_heatmap(grocery_value, coord_system = "calendar") |> 
  e_visual_map(max = max_value ) |> 
  e_title("Calendar", "Heatmap")|>
  e_tooltip(c("item", "axis")) 
}

calendar_plot(data_calendar, 2024)
# calendar_plot(data_calendar, 2023) 
```

### Supermarket

Value spent by supermarket

```{r}
supermakets_choosen <- c('Tesco', 'Lidl', 'Aldi')

grocery |>
  filter(supermarket_name %in% supermakets_choosen) |>
  group_by(grocery_day, supermarket_name, Year) |>
  summarise(grocery_value = value_total |> 
              sum(na.rm =  TRUE)) |>
  filter(grocery_value > 0) |>
  group_by(supermarket_name) |>
    e_charts(grocery_day) |> 
  e_line(grocery_value) |> 
  e_title("Groceries per supermarket")  |>
  e_tooltip(trigger = "axis")
```

Average spent by supermarket by Year

```{r}
grocery_supermarket_grouped <- grocery |>
  group_by(grocery_day, supermarket_name, Year) |>
  summarise(grocery_value = value_total |> 
              sum(na.rm =  TRUE)) |>
  filter(grocery_value > 0) |>
  ungroup() |>
  group_by(supermarket_name, Year) |>
  summarise(grocery_avg = grocery_value |> mean(),
            grocery_total = grocery_value |> sum(),
            grocery_min = grocery_value |> min(),
            grocery_max = grocery_value |> max()) |>
  arrange(desc(grocery_total))

  grocery_supermarket_grouped |>
  group_by(Year) |>
   e_charts(supermarket_name) |> 
  e_polar() |> 
  e_angle_axis() |> 
  e_radius_axis(supermarket_name) |> 
  e_bar(grocery_avg, coord_system = "polar") |> 
  e_scatter(grocery_avg, coord_system = "polar") |>
    e_title("Groceries per supermarket average anual")  |>
  e_tooltip(trigger = "axis")
```

```{r}
grocery_supermarket_grouped |>
  group_by(Year) |>
    e_charts(supermarket_name) |> 
 e_bar(grocery_total) |> 
  e_title("Total yearly grocery by supermarket") |>
  e_tooltip(trigger = "axis")
```

Value spent by supermarket monthly

```{r}
grocery_total_month_supermarket <- grocery |>
  group_by(Year, Month, supermarket_name) |>
  summarise(grocery_total = value_total |> sum(na.rm =  TRUE)) |>
  filter(grocery_total >0) 

grocery_sunburst  <- grocery_total_month_supermarket |>
  ungroup() |>
  nest(data = c(Month, supermarket_name, grocery_total)) |>
  mutate(children = map(data, ~.x |>
          nest(data = c(supermarket_name, grocery_total)
               ) |>
          mutate(children = map(data, ~.x |>
                   rename(value = grocery_total) |>
                   select(name = supermarket_name, value)
                   )|>
                             rename(name = Month) |>
                               select(name, children)))) |>
  select(name = Year, children)

# Convert to list format required by echarts4r
grocery_sunburst_list <-  e_list(grocery_sunburst)

e_charts() |>
  e_sunburst(
    data = grocery_sunburst_list,
    radius = c("10%", "90%"),
    label = list(rotate = 'radial')
  ) |>
  e_title("Total Grocery Sales by Year, Month, and Supermarket")
```
